### 2023-02-03 definitions.tex
details

\alain{dire juste que SAFE = upward closed et BAD = downward closed.
envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
4 resilience problem (U,D), (U,U), (D,U), (D,D)}

@@

These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS are given as upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets SAFE and BAD.

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $SAFE =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $BAD =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of SAFE. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that BAD is always downward closed and SAFE is always upward closed.
%		
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

@@

To decide the UkRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer. if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $\pred^n(SAFE)=\pred^*(SAFE)$ and then to test whether $n \leq k$. S is uniformely $k$-resilient iff $n \leq k$ and then UkRP is decidable.
\end{proof}

URP is decidable for lossy counter machines (LCM) with SAFE and BAD semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.

>>>>

%
%\alain{dire juste que SAFE = upward closed et BAD = downward closed.
%envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
%4 resilience problem (U,D), (U,U), (D,U), (D,D)}

@@

\alain{trouver les bad et safe maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ? on pourrait décider que safe=complement de bad et donc on aurait un seul ensemble (et son complement)}

These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets SAFE and BAD.
\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $SAFE =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $BAD =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of SAFE. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that BAD is always downward closed and SAFE is always upward closed.
%		
URP is decidable for lossy counter machines (LCM) with SAFE and BAD semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition. All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.

\begin{proposition}\label{general}
$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient iff $bad \subseteq \pred^*(safe)$.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-$k$-resilient iff $bad \subseteq \pred^k(safe)$.
\end{proposition}

\begin{proof}
Resilient means that 
\end{proof}

\subsection{Case: $SAFE=\uparrow SAFE$.}

\begin{theorem}\label{up-down}
The three uniform resilience problems are decidable for WSTS with effective predbasis and $safe=\uparrow safe$ and bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient. Since $safe=\uparrow safe$, and WSTS, $\pred^*(safe)=\uparrow \pred^*(safe)$, so $\pred^*(safe)$ admits a finite basis $B_{\pred^*(safe)}$. since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(safe)$ with the backward coverability algorithm. 
%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(safe)}$ which remais to compare, in both cases, two finite sets. 
if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(safe)}$ such that $s \leq b$. similar reasonning for the other case.
hence the two uniform resilience problems are decidable.

Let us examine the uniform $k$-resilience problem (UkRP). we may compute the least $n$ such that $ \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$. S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
\end{proof}

%One begins to test whether $Bad \cap \pred^*(safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans safe.\\

If $Bad \cap \pred^*(safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(safe)$.
resilient iff $Bad \subseteq \pred^*(safe)$  (eq $Bad \cap \pred^*(safe) = Bad$)

$Bad \cap \pred^*(safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(safe)$ et c'est (Bad',safe)-resilient.

\subsection{Case: $BAD=\downarrow BAD$ and $SAFE=\uparrow SAFE$.}

%
%		cas étudié par les deux articles sur les graphes
%

@@

To decide the UkRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer. if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$. S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
\end{proof}

%

\subsection{Case: $BAD=\uparrow BAD$ and $SAFE=\uparrow SAFE$.}

%%

\subsection{Case: $BAD=\downarrow BAD$ and $SAFE=\downarrow SAFE$.}

%%%%%%%%
\subsection{Case: $BAD=\uparrow BAD$ and $SAFE=\downarrow SAFE$.}
%
%		cas exclusion mutuelle
%
je prends les minimaux de bad et les idéaux de safe, pas sur que ce soit decidable....$pred(safe)$
